---
title: "Introduction to R"
author: "Sarah LaRocca"
date: "24 June 2016"
output:
  revealjs::revealjs_presentation:
    theme: solarized
    highlight: tango
    center: false
    transition: slide
    reveal_options:
      slideNumber: true
---

## Outline

- Introduction
- Getting started
- R basics
- Manipulating data
- Modeling data
- Visualizing data
- Source control

## Introduction

- `R` is a software environment for data manipulation, calculation, modeling, and graphical display.
- Primary use case is for data analysis and statistics, but has many other potential applications.
- Consists of a base environment with a small number of standard packages, and many optional packages which can be installed as needed.
- Has amusing version names--the most recent version is called "Bug in Your Hair" and the previous version was called "Very, Very Secure Dishes."

Source: https://cran.r-project.org/doc/manuals/R-intro.pdf

## Getting started

- R and RStudio Server are installed on an edge node of our cluster and can be accessed via a web browser.
- Log in using your AZCORP ID and password at: http://rstudio.amer.zurich.corp:8787/

## R basics
```{r, eval=FALSE}

# Comments start with a hash symbol

q() # quit

help() # help

getwd() # get current working directory

setwd('/home/testdirectory/') # set working directory to testdirectory

rm(list=ls()) # remove all objects from the workspace

```

## Expressions and assignment

- Standard arithmetic operators: `+`, `-`, `*`, `/`

- Relational operators: `<=`, `<`, `==`, `>`, `>=`, `!=`

- Logical operators: `|`, `&`

- Assignment operator: `<-`

## Expressions and assignment
```{r}
1 + 2
1 + 2 == 3
x <- 3
x

```

## Basic data types

- Numeric
- Integer
- Complex
- Logical
- Character

`class(x)` will give the class name of `x`.

## Vectors

- A vector is a single entity consisting of an ordered collection of numbers.
```{r}
a <- c(1,2,5.3,6,-2,4) # numeric vector
b <- c("apple","banana","pear") # character vector
c <- c(TRUE,TRUE,TRUE,FALSE,TRUE,FALSE) #logical vector

```

## Arrays and matrices

- An array can be considered as a multiply subscripted collection of data entries, for example numeric.
- A matrix is a two-dimensional array.
```{r}
A = matrix( 
    c(2, 4, 3, 1, 5, 7), # the data elements 
    nrow=2,              # number of rows 
    ncol=3,              # number of columns 
    byrow = TRUE)        # fill matrix by rows

A
```

## Lists

- A list is an object consisting of an ordered collection of objects known as its components.
- Components do not need to be of the same type.
```{r}
person <- list(name="Fred",
               wife="Mary",
               no_children=3,
               child_ages=c(4,7,9))

person

```
## Lists

```{r}
n = c(2, 3, 5) 
s = c("aa", "bb", "cc", "dd", "ee") 
b = c(TRUE, FALSE, TRUE, FALSE, FALSE) 
x = list(n, s, b, 3)   # x contains copies of n, s, b

x

```

## Data frames

- A data frame is a list of vectors with the same length.
- The components must be vectors (numeric, character, or logical), factors, numeric matrices,
lists, or other data frames.

```{r}
n = c(2, 3, 5) 
s = c("aa", "bb", "cc") 
b = c(TRUE, FALSE, TRUE) 
df = data.frame(n, s, b) 

df
```

## Useful `R` libraries from Hadley Wickham

- `dplyr` shortcuts for subsetting, summarizing, rearranging, and joining together data sets
- `tidyr` tools for changing the layout of your data sets
- `stringr` tools for regular expressions and character strings
- `lubridate` tools for working with dates and times
- `ggplot2` uses the 'grammar of graphics' to build layered, customizable plots

```{r, eval=FALSE}
library(dplyr)
```

## Reading data
```{r}
weather <- read.csv('data/weather.csv')

class(weather)
summary(weather)
```
